---
title: "Processed code"
author: "Vicky Jia"
date: "4/17/2020"
output:
  word_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---
```{r}
#Set up
getwd()
library(tidyverse)
library(cowplot)
library(lubridate)
library(zoo)
library(knitr)
library(kableExtra)
library(viridis)
library(agricolae)
library(corrplot)
library(trend)
library(modifiedmk)
projecttheme <- theme_light() + 
  theme(
    plot.title = element_text(color = 'dimgray', size = 12, face = 'bold'), 
    axis.text = element_text(color = "black", size= 10), 
    axis.title = element_text(color = 'black', size = 11, face = 'bold'),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.text = element_text(size = 10),
    legend.position = "bottom")
theme_set(projecttheme)
```

#GDP
```{r}
GDP <- read.csv('./Raw Data/BEA_GDP_raw.csv')
str(GDP)
colnames(GDP)
GDP <- GDP %>%
  rename('GDP' = 'Gross.domestic.product', 'Personal_consumption' = 'Personal.consumption.expenditures')

GDP.select<- GDP %>%
  select(Year, GDP, Personal_consumption) %>%
  filter(Year>1989 & Year < 2018)
dim(GDP.select)

kable(head(GDP.select), caption = "First 6 years of GDP and personal consumption (billion US$)") %>%
  kable_styling(latex_options="scale_down")

GDP.gather <- gather(GDP.select, 'Type', 'Value', -Year)
count(GDP.gather)

GDP.line <- ggplot(GDP.gather)+
  geom_line(aes(x= Year, y=Value, color = Type))+
  labs(y= 'US$ (billion)',
       title = 'US Economic Features in 1990-2017',
       color = '')+
  scale_color_brewer(type='qual', direction = -1)+
  scale_x_continuous("Year", labels = as.character(GDP.gather$Year), breaks = GDP.gather$Year)
print(GDP.line)

write.csv(GDP.select, row.names = FALSE, file = './Processed Data/BEA_GDP_processed.csv')
```

#Electricity Consumption
```{r}
electric <- read.csv('./Raw Data/EIA_electricity-consumption_sector_raw.csv')

str(electric)
colnames(electric)
electric <- electric %>%
  rename('Industry_Sector' = Industry.Sector.Category, 'elec_residential' = Residential, 'elec_commercial' = Commercial, 'elec_industrial' = Industrial, 'elec_transport' =Transportation)
colnames(electric)

elec.select <- electric %>%
  filter(State == 'US'& Year>1989 & Year < 2018 & Industry_Sector == 'Total Electric Industry')%>%
  select(Year, elec_residential, elec_commercial, elec_industrial, elec_transport) 
dim(elec.select)

kable(head(elec.select), caption = "First 6 years of electricity consumptions by sector in US (Mwh)") %>%
  kable_styling(latex_options="scale_down")

elec.gather <- gather(elec.select, 'Sector', 'Electricity_Consumption', -Year)
count(GDP.gather)
elec.gather$Electricity_Consumption <- str_replace_all(elec.gather$Electricity_Consumption,
pattern = "([,])", replacement = "")

elec.gather$Electricity_Consumption <- as.numeric(elec.gather$Electricity_Consumption)
str(elec.gather)
elec.gather2 <- elec.gather %>%
  filter(!is.na(Electricity_Consumption))

elec.line <- ggplot(elec.gather2)+
  geom_line(aes(x= Year, y=Electricity_Consumption, color=Sector))+
  labs(y= 'Electricity Consumption (Mwh)',
       title = 'US Electricty Consumption by Sector in 1990-2017',
       color = '')+
  scale_color_brewer(palette = 'Set1')+
  #scale_x_continuous("Year", labels = as.character(elec.gather2$Year), breaks = elec.gather2$Year)+
  facet_wrap(vars(Sector), scales = 'free')+
  theme(legend.position = "none") 
print(elec.line)

write.csv(elec.select, row.names = FALSE, file = './Processed Data/EIA_electricity-consumption_sector_processed.csv')

```

#GHG gases
```{r}
GHG.gas <- read.csv('./Raw Data/EPA_GHG_Gas_raw.csv')

str(GHG.gas)
colnames(GHG.gas)
GHG.gas <- GHG.gas %>%
  rename("CO2" = Carbon.dioxide, 'CH4' = Methane, 'N2O' = Nitrous.oxide, 'Fgas' = Fluorinated.gases)
colnames(GHG.gas)

kable(head(GHG.gas), caption = "First 6 years of GHG emissions by gas type (MMTCO2e)") %>%
  kable_styling(latex_options="scale_down")

GHG.gas.gather <- gather(GHG.gas, 'Gas', 'Concentration', -Year)
count(GHG.gas.gather)

GHG.gas.line <- ggplot(GHG.gas.gather)+
  geom_line(aes(x= Year, y=Concentration, color = Gas))+
  labs(y= bquote(bold('Concentration (MMT'~CO[2]~'e)')),
       title = bquote(bold('US GHG Emissions by Gas Type in 1990-2017 (MMT'~CO[2]~'e)')),
       color = '')+
  scale_color_brewer(palette = 'Set1')+
  scale_x_continuous("Year", labels = as.character(GHG.gas.gather$Year), breaks = GHG.gas.gather$Year)

GHG.gas.line.facet <- ggplot(GHG.gas.gather)+
  geom_line(aes(x= Year, y=Concentration, color = Gas))+
  labs(y= bquote(bold('Concentration (MMT'~CO[2]~'e)')),
       color = '')+
  scale_color_brewer(palette = 'Set1')+
  facet_wrap(vars(Gas), scales = 'free')+
  theme(legend.position = "none")

plot_grid(GHG.gas.line,
          GHG.gas.line.facet,
          label_x = 0.2,
          nrow = 2)

write.csv(GHG.gas, row.names = FALSE, file = './Processed Data/EPA_GHG_Gas_processed.csv')
```

#GHG sector
```{r}
GHG.sector <- read.csv('./Raw Data/EPA_GHG_Sector_raw.csv')

str(GHG.sector)
colnames(GHG.sector)
GHG.sector <- GHG.sector %>%
  rename('Electricity_generation' = Electricity.generation)
  #rename('ghg_elecgene' = Electricity.generation, 'ghg_transport' =Transportation, 'ghg_industry'=Industry, 'ghg_agri' =Agriculture, 'ghg_commercial' =Commercial, 'ghg_residential' =Residential)
colnames(GHG.sector)

GHG.sector.select <- GHG.sector %>%
  select(Year, Transportation, Electricity_generation, Industry, Agriculture, Commercial,Residential)
  #select(Year, ghg_transport, ghg_elecgene, ghg_industry, ghg_agri, ghg_commercial, ghg_residential)
dim(GHG.sector.select)

kable(head(GHG.sector.select), caption = "First 6 years of GHG emissions by sector (MMTCO2e)") %>%
  kable_styling(latex_options="scale_down")

GHG.sector.gather <- gather(GHG.sector.select, 'Sector', 'Emission', -Year)
count(GHG.sector.gather)

GHG.sector.line <- ggplot(GHG.sector.gather)+
  geom_line(aes(x= Year, y=Emission, color = Sector))+
  labs(y= bquote(bold('GHG Emission (MMT'~CO[2]~'e)')),
       title = bquote(bold('US GHG Emissions by Gas Type in 1990-2017 (MMT'~CO[2]~'e)')),
       color = '')+
  scale_color_brewer(palette = 'Dark2')+
  scale_x_continuous("Year", labels = as.character(GHG.sector.gather$Year), breaks = GHG.sector.gather$Year)
plot(GHG.sector.line)

GHG.sector.line.facet <- ggplot(GHG.sector.gather)+
  geom_line(aes(x= Year, y=Emission, color = Sector))+
  labs(y= bquote(bold('GHG Emission (MMT'~CO[2]~'e)')),
       color = '')+
  scale_color_brewer(palette = 'Dark2')+
  facet_wrap(vars(Sector), scales = 'free')+
  theme(legend.position = "none")

plot_grid(GHG.sector.line,
          GHG.sector.line.facet,
          label_x = 0.2,
          nrow = 2)
write.csv(GHG.sector.select, row.names = FALSE, file = './Processed Data/EPA_GHG_Sector_processed.csv')
```

#Temperature
```{r}
temp <- read.csv('./Raw Data/NOAA_temp_raw.csv')
str(temp)
colnames(temp)
temp <- mutate(temp, Year = year(as.Date(as.yearmon(as.character(temp$Date), format = "%Y%m"))))
str(temp)

temp.select <- temp %>%
  select(Year, Value, Anomaly) %>%
  mutate(Mean_all_temp =  mean(Value))
dim(temp.select)
str(temp.select)

kable(head(temp.select), caption = "First 6 years of US temperature data in 1990-2017 (F)") %>%
  kable_styling(latex_options="scale_down")

temp.line <- ggplot(temp.select, aes(x= Year, y=Value))+
  geom_point()+
  geom_line()+
  geom_line(aes(y=Mean_all_temp, color= 'red'))+
  labs(y= expression("Temperature ("*~degree*F*")"),
       title = expression("US Annual Mean Temperature in 1990-2017 ("*~degree*F*")"))+
  scale_x_continuous("Year", labels = as.character(temp.select$Year), breaks = temp.select$Year)+
  geom_text(x = 2017, y = 53, label = "Average Temperature for 18 Years", hjust = 1, size = 3, color = 'red')+
  theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_blank())
plot(temp.line)

temp.select$colour <- ifelse(temp.select$Anomaly < 0, "negative","positive")
temp.anom <- ggplot(temp.select)+
  geom_col(aes(x=Year, y=Anomaly, fill=colour))+
  labs(y= expression("Temperature Anomaly ("*~degree*F*")"),
      title = expression("US Annual Temperature Anomaly in 1990-2017 ("*~degree*F*")"),
      color = '')+
  scale_x_continuous("Year", labels = as.character(temp.select$Year), breaks = temp.select$Year)+
  scale_fill_manual(values=c(positive="firebrick1",negative="steelblue"))+
  theme(legend.position = 'none')
plot(temp.anom)

plot_grid(temp.line,
          temp.anom,
          label_x = 0.2,
          nrow = 2)

write.csv(temp.select, row.names = FALSE, file = './Processed Data/NOAA_temp_processed.csv')
```

#Population
```{r}
ppl <- read.csv('./Raw Data/WB_pop_raw.csv')

str(ppl)
ppl.select <- ppl %>%
  select(Year, United.States) %>%
  filter(Year >1989 & Year < 2018)
dim(ppl.select)

colnames(ppl.select)
ppl.select <- ppl.select %>%
  rename('US_pop' = United.States)
colnames(ppl.select)

kable(head(ppl.select), caption = "First 6 years of US population in 1990-2017") %>%
  kable_styling(latex_options="scale_down")


write.csv(ppl.select, row.names = FALSE, file = './Processed Data/WB_pop_processed.csv')
```

#Analysis part - rearrange dataset
```{r}
newGDP <- read.csv('./Processed Data/BEA_GDP_processed.csv', head(T))
newelectricity <- read.csv('./Processed Data/EIA_electricity-consumption_sector_processed.csv', head(T))
newGHGgas <- read.csv('./Processed Data/EPA_GHG_Gas_processed.csv', head(T))
newGHGsector <- read.csv('./Processed Data/EPA_GHG_Sector_processed.csv', head(T))
newtemp <- read.csv('./Processed Data/NOAA_temp_processed.csv', head(T))
newpop <- read.csv('./Processed Data/WB_pop_processed.csv', head(T))

allGHG <- full_join(newGHGgas,newGHGsector, by = 'Year')
GHG_contribute <- full_join(newelectricity, newpop, by = 'Year')
GHG_impact <- full_join(newGDP, newtemp, by = 'Year')
```

#Analysis part - statistical tests
lm - look at explainations here https://feliperego.github.io/blog/2015/10/23/Interpreting-Model-Output-In-R
##1. normal distrbution GHG concentration
shapiro test: two p-values are greater than the significance level 0.05 implying that the distribution of the data are not significantly different from normal distribution. In other words, we can assume the normality.
```{r}
summary(allGHG)

shapiro.test(allGHG$Total)
shapiro.test(allGHG$CO2)
shapiro.test(allGHG$CH4)
shapiro.test(allGHG$N2O)
shapiro.test(allGHG$Fgas)

shapiro.test(allGHG$Transportation)
shapiro.test(allGHG$Electricity_generation)
shapiro.test(allGHG$Industry)
shapiro.test(allGHG$Agriculture)
shapiro.test(allGHG$Commercial)
shapiro.test(allGHG$Residential)

ggplot(allGHG, aes(x=Total))+
  geom_histogram()
qqnorm(allGHG$Total);qqline(allGHG$Total)

allsample <- t.test(allGHG, mu=mean(allGHG$Total), alternative = 'less')
allsample

```

##2. Does total electricity consumption differ among sectors (category vs discrete)?
Shapiro-Wilk test can be performed as follow:
Null hypothesis: the data are normally distributed
Alternative hypothesis: the data are not normally distributed

lm: If the p-value is less than 0.05, we reject the null hypothesis that there's no difference between the means and conclude that a significant difference does exist.
```{r}
newelectricity.gather <- gather(newelectricity, 'Sector', 'Emission', -Year)
count(newelectricity.gather)

newelectricity.gather$Emission <- str_replace_all(newelectricity.gather$Emission,
pattern = "([,])", replacement = "")
newelectricity.gather$Emission <- as.numeric(newelectricity.gather$Emission)
str(newelectricity.gather)
#elec.gather2 <- elec.gather %>%
 # filter(!is.na(Electricity_Consumption))

#One-way Anova, test for normality
shapiro.test(newelectricity.gather$Emission[newelectricity.gather$Sector=='elec_residential'])

shapiro.test(newelectricity.gather$Emission[newelectricity.gather$Sector=='elec_commercial'])

shapiro.test(newelectricity.gather$Emission[newelectricity.gather$Sector=='elec_industrial'])

shapiro.test(newelectricity.gather$Emission[newelectricity.gather$Sector=='elec_transport'])

qqnorm(newelectricity.gather$Emission); qqline(newelectricity.gather$Emission)

#Test for equal variance
##a small p-value for the intercept and the slope indicates that we can reject the null hypothesis which allows us to conclude that there is a relationship between speed and distance.
bartlett.test(newelectricity.gather$Emission ~ newelectricity.gather$Sector)


newelectricity.anova <- aov(data = newelectricity.gather, Emission ~ Sector)
summary(newelectricity.anova)

#Checking model fit and assumptions
# ANOVA is robust against departures from normality.
par(mfrow = c(2,2), mar=c(1,1,1,1))
plot(newelectricity.anova)

# Post-hoc test # Run a post-hoc test for pairwise differences
TukeyHSD(newelectricity.anova)

#Extract groupings for pairwise relationships
newelectricity.groups <- HSD.test(newelectricity.anova, "Sector", group = TRUE)
newelectricity.groups

newelectricity.graph <- ggplot(newelectricity.gather, aes(x = Sector, y = Emission)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + stat_summary(geom = "text", fun.y = max, vjust = -1, size = 3.5, label = c('b','c','a','d')) +
  labs(x = "Sector", y = "Emission") 
print(newelectricity.graph)


```

##Move up--3. Are sectoral GHG emissions significantly different from each other? How sectoral emissions contribute to total emissions (correlation)? (Lec12 - multiple regression)
cor.test: If the p-value is < 5%, then the correlation between x and y is significant.
```{r}
##sectoral emissions differ from each other?
GHG.regression <- lm(data=allGHG, Transportation~ Electricity_generation+Industry+Agriculture+Commercial+Residential)
summary(GHG.regression)
step(GHG.regression)

#how contribute to total emission - correlation
cor.test(allGHG$Total,allGHG$Transportation)
cor.test(allGHG$Total,allGHG$Electricity_generation)
cor.test(allGHG$Total,allGHG$Industry)
cor.test(allGHG$Total,allGHG$Agriculture)
cor.test(allGHG$Total,allGHG$Commercial)
cor.test(allGHG$Total,allGHG$Residential)

par(mfrow = c(2,2), mar=c(1,1,1,1))
plot(GHG.regression)

par(mfrow = c(1,1))
allGHG.bysector <- allGHG %>%
  select(Total:Residential)
allGHG.corr <- cor(allGHG.bysector)
corrplot(allGHG.corr, method = "ellipse", tl.cex = 0.5)
```


##4. Are electricity comsupmtions in some sectors significant to total GHG emissions?
lm: If the p-value is less than 0.05, we reject the null hypothesis which allows us to conclude that there is a relationship between speed and distance.
```{r}
##linear regression - electricity consumption relation to total GHG emissions
GHG_contribute$elec_residential <- str_replace_all(GHG_contribute$elec_residential,
pattern = "([,])", replacement = "")
GHG_contribute$elec_residential <- as.numeric(GHG_contribute$elec_residential)
GHG_contribute$elec_commercial <- str_replace_all(GHG_contribute$elec_commercial,
pattern = "([,])", replacement = "")
GHG_contribute$elec_commercial <- as.numeric(GHG_contribute$elec_commercial)
GHG_contribute$elec_industrial <- str_replace_all(GHG_contribute$elec_industrial,
pattern = "([,])", replacement = "")
GHG_contribute$elec_industrial <- as.numeric(GHG_contribute$elec_industrial)
GHG_contribute$elec_transport <- str_replace_all(GHG_contribute$elec_transport,
pattern = "([,])", replacement = "")
GHG_contribute$elec_transport <- as.numeric(GHG_contribute$elec_transport)

par(mfrow = c(2,2), mar=c(1,1,1,1))
elec.regression <- lm(allGHG$Total ~GHG_contribute$elec_residential + GHG_contribute$elec_commercial + GHG_contribute$elec_industrial + GHG_contribute$elec_transport)
summary(elec.regression)
plot(elec.regression)

step(elec.regression)

elecmodel <- lm(formula = allGHG$Total ~ GHG_contribute$elec_residential + 
    GHG_contribute$elec_transport)
summary(elecmodel)
plot(elecmodel)


```

##5. Is population growth significant to GHG emissions ?
cor.test: If the p-value is < 5%, then the correlation between x and y is significant.
```{r}
shapiro.test(GHG_contribute$US_pop) #normal

pop.regression <- lm(allGHG$Total ~GHG_contribute$US_pop)
summary(pop.regression)
cor.test(allGHG$Total,GHG_contribute$US_pop) #not significant, no relationship, same as lm
plot(pop.regression)

par(mfrow = c(1,1))
ppl.all <- full_join(allGHG, newpop, by = 'Year')
# Plot the regression
GHGpop.plot <-ggplot(ppl.all, aes(x = US_pop, y = Total)) + 
  geom_line()+
  labs(y= bquote(bold('GHG Emission (MMT'~CO[2]~'e)')),
       x= 'Population',
       color = '')+
  geom_vline(xintercept = ppl.all$US_pop[which.max(ppl.all$Total)], color = "#f20505", lty = 2, size = 0.9)+
  geom_text(y = 6400, x = 3.0e+08, label = 'Highest GHG Concentration Year: 2007', hjust = 1, size = 3, color = 'red')
print(GHGpop.plot)

```

#Part 2 - impacts
##Move up--1. Are emissions of different types of GHGs related to each other?
lm: a small p-value for the intercept and the slope indicates that we can reject the null hypothesis which allows us to conclude that there is a relationship between speed and distance.
```{r}
GHG.gas.regression <- lm(data=allGHG, CO2~ CH4+N2O+Fgas)
summary(GHG.gas.regression)
step(GHG.gas.regression)

par(mfrow = c(2,2), mar=c(1,1,1,1))
plot(GHG.gas.regression)

GHG.gas.regression.improve <- lm(formula = CO2 ~ CH4 + N2O, data = allGHG)
summary(GHG.gas.regression.improve)
plot(GHG.gas.regression.improve)

par(mfrow = c(1,1))

allGHG.bygas <- allGHG %>%
  select(CO2:Fgas)

cor.test(allGHG.bygas$CO2, allGHG.bygas$CH4)
cor.test(allGHG.bygas$CO2, allGHG.bygas$N2O)
cor.test(allGHG.bygas$CO2, allGHG.bygas$Fgas)
#Visualize this result
allGHG.gas.corr <- cor(allGHG.bygas)
corrplot(allGHG.gas.corr, method = "ellipse", tl.cex = 0.5)
```

##2. GHGs and temperature 
which sector emission is significant to temperature anomaly?
```{r}
shapiro.test(newtemp$Anomaly)
#time series analysis- Mann-Kendall
#The null hypothesis of the test is that there is no trend in the data and the alternative hypothesis is that the data represents a monotonic trend.

#If the p-value is less than 0.05, we reject the null hypothesis.
GHGtemp_data<- full_join(allGHG, newtemp, by='Year')
GHGtemp.ts <- ts(GHGtemp_data$Value, start=c(1990), end=c(2017), frequency=1)
class(GHGtemp.ts)
GHGtemp.ts.vector <- as.numeric(GHGtemp.ts)
class(GHGtemp.ts.vector)
GHGtemp.trend <- mmkh(GHGtemp.ts.vector, ci=0.95)
GHGtemp.trend
summary(GHGtemp.trend)


#linear regression
##a p-value of 5% or less is a good cut-off point. In our model example, the p-values are very close to zero. Note the ‘signif. Codes’ associated to each estimate. Three stars (or asterisks) represent a highly significant p-value. Consequently, a small p-value for the intercept and the slope indicates that we can reject the null hypothesis which allows us to conclude that there is a relationship between speed and distance.
GHG.temp.regression <- lm(GHGtemp_data,formula = Anomaly~  Transportation+Electricity_generation+Industry+Agriculture+Commercial+Residential)
summary(GHG.temp.regression)

par(mfrow = c(2,2), mar=c(1,1,1,1))
plot(GHG.temp.regression)

step(GHG.temp.regression)

GHG.temp.regression.improve <- lm(formula = Anomaly ~ Transportation + Agriculture + Commercial + Residential, data = GHGtemp_data)
summary(GHG.temp.regression.improve)

plot(GHG.temp.regression.improve)
par(mfrow = c(1,1))

GHGsector.temp <- GHGtemp_data %>%
  select(Transportation:Residential, Anomaly)

cor.test(GHGsector.temp$Anomaly, GHGsector.temp$Transportation)
cor.test(GHGsector.temp$Anomaly, GHGsector.temp$Electricity_generation)
cor.test(GHGsector.temp$Anomaly, GHGsector.temp$Industry)
cor.test(GHGsector.temp$Anomaly, GHGsector.temp$Agriculture)
cor.test(GHGsector.temp$Anomaly, GHGsector.temp$Commercial)
cor.test(GHGsector.temp$Anomaly, GHGsector.temp$Residential)

GHGsector.temp.corr <- cor(GHGsector.temp)
corrplot(GHGsector.temp.corr, method = "ellipse", tl.cex = 0.5)
```

##3. GHG and GDP
Is GHG emission significant to GDP growth?
```{r}
shapiro.test(newGDP$GDP)

GHGgdp_data<- full_join(allGHG, newGDP, by='Year')
GHGgdp.ts <- ts(GHGgdp_data$GDP, start=c(1990), end=c(2017), frequency=1)
class(GHGgdp.ts)
GHGgdp.ts.vector <- as.numeric(GHGgdp.ts)
class(GHGgdp.ts.vector)
GHGgdp.trend <- mmkh(GHGgdp.ts.vector, ci=0.95)
GHGgdp.trend
summary(GHGtemp.trend)

GDP.anova <- lm(data=GHGgdp_data, Total~GDP+Personal_consumption)
summary(GDP.anova) 

par(mfrow = c(2,2), mar=c(1,1,1,1))
plot(GDP.anova)

par(mfrow = c(1,1))

step(GDP.anova)

# Add a line and standard error for the linear regression
GHGgdp.plot <- ggplot(GHGgdp_data, aes(x = Total, y = GDP)) + 
  geom_smooth(method = "lm") +
  labs(y= 'GDP (US$ billion)', x= bquote(bold('Total GHG Emission (MMT'~CO[2]~'e)')))+
  geom_point()
print(GHGgdp.plot)
```

