---
output: 
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
geometry: margin=2.54cm
title: "Factors and Impacts of Greenhouse Gas Emissiosn in US from 1990-2017"
subtitle: "Web address for GitHub repository"
author: "Vicky Jia"
fontsize: 12pt
mainfont: Times New Roman
editor_options: 
  chunk_output_type: console
---

\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

```{r setup, include=FALSE}
setwd("~/DataAnalytics_Final")
library(tidyverse)
library(cowplot)
library(lubridate)
library(zoo)
library(knitr)
library(kableExtra)
library(viridis)
library(agricolae)
library(corrplot)
library(trend)
library(modifiedmk)

projecttheme <- theme_light() + 
  theme(
    plot.title = element_text(color = 'dimgray', size = 12, face = 'bold'), 
    axis.text = element_text(color = "black", size= 10), 
    axis.title = element_text(color = 'black', size = 11, face = 'bold'),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.text = element_text(size = 10),
    legend.position = "bottom")
theme_set(projecttheme)

GDP <- read.csv('./Raw Data/BEA_GDP_raw.csv')
electric <- read.csv('./Raw Data/EIA_electricity-consumption_sector_raw.csv')
GHG.gas <- read.csv('./Raw Data/EPA_GHG_Gas_raw.csv')
GHG.sector <- read.csv('./Raw Data/EPA_GHG_Sector_raw.csv')
temp <- read.csv('./Raw Data/NOAA_temp_raw.csv')
ppl <- read.csv('./Raw Data/WB_pop_raw.csv')
structure <- read.csv('./Processed Data/Processed_Data_Structure.csv')

```


#Rationale and Research Questions

Global warming resulted from greenhouse gas (GHG) emissions is one of the biggest challenges to human beings. While IPCC in its latest special report warns people the critial consequences of 1.5-degree-Celcius temperature increase from the pre-industrial level, the average global temperature on Earth has already increased by about 0.8° Celsius (1.4° Fahrenheit) since 1880. 

In the United States, in the period of 1990-2017, the temperature increase and GHG emissions have been greatly contribetd by the energy sector. This was attributed to the increase electricity consumptions in all sectors by more people, and electricity generation primarily relies on fossil-fuel combustions. In addition to temperature increase, the economic features also have significant changes along the increasing GHG emissions and electricity consumption in the meantime. Understanding environmental and socioeconomic impacts of GHG emissions is significant to policy makers to consider long-term decision making. Therefore, this project examines the impacts of energy consumptions and demographic changes on GHG emissions, as well as further investigates the environmental and economic effects of GHG emissions in the United States from 1990 to 2017. 

Specifically, this project aims to study factors that might contribute to GHG emissions guided by the following questions:

1. What were impacts of US electricity consumption changes on GHG emissions from 1990 to 2017? What sectors of electricity econsumption were significant to GHG emissions in these 18 years?
2. Was the population growth a signifant factor in affecting GHG emissisions in this period?

Following up with GHG emission causes, this project will then explore the impacts of GHG emissions in US from 1990 to 2017 guided by the following questions: 

3. Which sectors were more significant to temperature anomalies in this period?
4. Was the economic development in this period, namely gross domestic product (GDP) and personal consumption expenditures, significantly realted to the GHG emissions? 

\newpage

#Dataset Information

## Raw Database Information
There were 6 datasets used in this project retrieved from various websites. Detailed criteria to collect these raw data in csv file format are listed below. 

### BEA_GDP_raw.csv:
US annual gross domestic production and personal consumption expenditures data from 1990-2017.Data were retrieved from US Bureau of Economic Analysis National Income and Product Accounts (NIPA) Interactive Data Tables tool. Data were transposed by the investigator.       
(https://apps.bea.gov/iTable/iTable.cfm?reqid=19&step=2#reqid=19&step=2&isuri=1&1921=survey accessed on 2020-04-11.)

The following selections were made:       
*First Year: 1990-A&Q       
*Last Year: 2017-A&Q        
*Scale: billion           
*Series: Annual


### EIA_electricity-consumption_sector_raw.csv:
The annual retail sales of electricity to ultimate customers by sector, by state, by provider from 1990 to 2018 in US. The data was retrieved from Annual Electric Power Industry Report, Form EIA-861 detailed data files from US Energy Information Administration.

(https://www.eia.gov/electricity/data/state/sales_annual.xlsx accessed on 2020-04-11.)


### EPA_GHG_Gas_raw.csv:
US GHG emission data by gas types in all sectors from 1990-2017 provided by EPA's annual Inventory of U.S. Greenhouse Gas Emissions and Sinks. Data was retrieved from US EPA Greenhouse Gas Inventory Data Explorer. Data were transposed by the investigator.       
(https://cfpub.epa.gov/ghgdata/inventoryexplorer/#allsectors/allgas/gas/all accessed on 2020-04-11.)

The following selections were made:         
*Sector: All sectors          
*Greenhouse gas: All gases        
*Break out by: Gas        
*Year: All years


### EPA_GHG_Sector_raw.csv:
GHG emission data by economic sectors for all greenhouse gases from 1990-2017 provided by EPA's annual Inventory of U.S. Greenhouse Gas Emissions and Sinks. Data was retrieved from US EPA Greenhouse Gas Inventory Data Explorer. Data were transposed by the investigator.        
(https://cfpub.epa.gov/ghgdata/inventoryexplorer/#allsectors/allgas/econsect/all accessed on 2020-04-11.)

The following selections were made: 
*Sector: All sectors          
*Greenhouse gas: All gases        
*Break out by: Economic sector        
*Year: All years


### NOAA_temp_raw.csv:
US mean annual temperature and temperature anomalies from 1990 to 2017. Data was retrieved from NOAA National Centers for Environmental Information Climate at a Glance.       
(https://www.ncdc.noaa.gov/cag/national/time-series/110/tavg/ann/12/1990-2017?base_prd=true&begbaseyear=1901&endbaseyear=2000, accessed on 2020-04-11.)

The following selections were made:       
*Parameter: Average Temperature       
*Time Scale: Annual         
*Start Year: 1990         
*End Year: 2017         
*Display Base Period: Start: 1990, End: 2017        


### WB_pop_raw.csv:
Total annual population data from 1960-2018 retrieved from World Bank dataset. Data were transposed by the investigator.       
(https://data.worldbank.org/indicator/SP.POP.TOTL?locations=US accessed on 2020-04-11.)

The following selections were made in the searching window:       
*Population, total          
*United States

## Data Structure
The raw data collected from some databases are beyond the scope of this project, so these datasets were wrangled to fit to the objective of this project. To provide of an overview of the data structure of each raw dataset, raw data were reorganized by gathering same types of data to one column together. Table 1 summarzies the data structure for all raw datasets used in this project after being gathered These datasets are saved in "Processed Data" folder.

The specific content of each dataset and the detailed process of reframing each raw dataset to meet the objective of this project are introduced in Chapter 3, Exploratory Analysis. 
```{r echo = FALSE, message = FALSE, warning = FALSE}
GDPraw.gather <- gather(GDP, 'Type', 'Value', -Year)
write.csv(GDPraw.gather, row.names = FALSE, file = './Processed Data/BEA_GDP_structure.csv')
elecraw.gather <- gather(electric, 'Sector', 'Electricity_Consumption', Residential:Total)
write.csv(elecraw.gather, row.names = FALSE, file = './Processed Data/EIA_electricity-consumption_sector_structure.csv')
GHG.gas.raw.gather <- gather(GHG.gas, 'Gas', 'Concentration', -Year)
write.csv(GHG.gas.raw.gather, row.names = FALSE, file = './Processed Data/EPA_GHG_Gas_structure.csv')
GHG.sector.raw.gather <- gather(GHG.sector, 'Sector', 'Emission', -Year)
write.csv(GHG.sector.raw.gather, row.names = FALSE, file = './Processed Data/EPA_GHG_Sector_structure.csv')
write.csv(temp, row.names = FALSE, file = './Processed Data/NOAA_temp_structure.csv')
ppl.raw.gather <- gather(ppl, 'Country', 'Population', -Year)
write.csv(ppl.raw.gather, row.names = FALSE, file = './Processed Data/WB_pop_structure.csv')

kable(structure, caption = "Summary of Data Structure ") %>%
  kable_styling(latex_options="scale_down")
```

\newpage

# Exploratory Analysis 
Some raw datasets are at global scale, or have longer temporal ranges, which are beyond the scope of this project. Also, some discrete variables were downloaded as characters, which could not be adopted by some statistical models. Therefore, thse datasets were first wrangled to focus on US region in the period between 1990 and 2017. This section demonsrates the process of data wrangling in details, as well as presents the overview of the content of each dataset.

##US GDP
```{r warning = FALSE}
#Show the structure of the raw dataset
str(GDP)

#Rename column name for easier use
GDP <- GDP %>%
  rename('GDP' = 'Gross.domestic.product', 'Personal_consumption' = 'Personal.consumption.expenditures')

##Reframe dataset to meet project objective
GDP.select<- GDP %>%
  filter(Year>1989 & Year < 2018) %>%
  select(Year, GDP, Personal_consumption)
dim(GDP.select)

```

Figure 1 illustartes the trend of change of US GDP and personal consumptions over the years. Generally, GDP growth was increasing in most years, though slowing down at the beginning of the 21st century and decreasing in 2008-2009 when financial crisis ocurred. The changing trend of personal consumptions over the years was generally consistent to GDP growth, as the personal consumption is one of variables accounted to total GDP. Therefore, from this overview of the dataset, the variable of personal consumption would not be considered as an unique and significant variable to conduct additional analysis. Instead, in economic development analysis in Chapter 4, only total GDP will be studied.

```{r echo = FALSE, message = FALSE, warning = FALSE, fig.cap= 'US Economic Features in 1990-2017'}
#kable(head(GDP.select), caption = "First 6 rows of GDP and personal consumption (billion US$)") %>%
 # kable_styling(latex_options="hold_position")

#Gather dataset together for visualization
GDP.gather <- gather(GDP.select, 'Type', 'Value', -Year)
count(GDP.gather)

GDP.line <- ggplot(GDP.gather)+
  geom_line(aes(x= Year, y=Value, color = Type))+
  labs(y= 'US$ (billion)',
       color = '')+
  scale_color_brewer(type='qual', direction = -1)+
  scale_x_continuous("Year", labels = as.character(GDP.gather$Year), breaks = GDP.gather$Year)
print(GDP.line)
```

\newpage

##US Electricity Consumption
There are three types of industry sector in the raw dataset. In this project, I focused on "Total Electric Industry", which includes all types of survices in electricity consumption in each sector. There are four sectors covered in the electricity consumption dataset, which are illustrateed as Table ???. The values of consumption data are in factor format and separted by commas, so it is necessary to convert these values to numeric variables first. By visualizing electricity consumption data in Figure 2, we can notice that all four graphs are in different consumption scales. Notably, transportation sector has the lowest electricity consumption among all sectors, and also there was no data available for transportation sector in before 2002. Therefore, transporation sector will be filtered out in the next step statistic analysis.

```{r warning = FALSE}
#Show the structure of the raw dataset
str(electric)

#Rename column name for easier use
electric <- electric %>%
  rename('Industry_Sector' = Industry.Sector.Category, 'elec_residential' = Residential, 'elec_commercial' = Commercial, 'elec_industrial' = Industrial, 'elec_transport' =Transportation)
colnames(electric)

##Reframe dataset to meet project objective
elec.select <- electric %>%
  filter(State == 'US'& Year>1989 & Year < 2018 & Industry_Sector == 'Total Electric Industry')%>%
  select(Year, elec_residential, elec_commercial, elec_industrial, elec_transport) 
dim(elec.select)
```

```{r echo = FALSE, message=FALSE, warning=FALSE, fig.cap='US Electricty Consumption by Sector in 1990-2017'}
#The preview of reframed dataset
kable(head(elec.select), caption = "First 6 rows of electricity consumptions by sector in US (Mwh)") %>%
  kable_styling(latex_options="hold_position")

#Gather dataset together for visualization
elec.gather <- gather(elec.select, 'Sector', 'Electricity_Consumption', -Year)
count(GDP.gather)

#Convert to numeric variables
elec.gather$Electricity_Consumption <- str_replace_all(elec.gather$Electricity_Consumption,
pattern = "([,])", replacement = "")
elec.gather$Electricity_Consumption <- as.numeric(elec.gather$Electricity_Consumption)
str(elec.gather)

#Remove NA data
elec.gather2 <- elec.gather %>%
  filter(!is.na(Electricity_Consumption))

elec.line <- ggplot(elec.gather2)+
  geom_line(aes(x= Year, y=Electricity_Consumption, color=Sector))+
  labs(y= 'Electricity Consumption (Mwh)',
       color = '')+
  scale_color_brewer(palette = 'Set1')+
  facet_wrap(vars(Sector), scales = 'free')+
  theme(legend.position = "none") 
print(elec.line)

```

\newpage

##US GHG Concentration by Gas
GHG concentration data indicates the concentration of different GHG gases emitted to the atmosphere. The upper graph in Figure 3 displays the trends for all GHGs together, from which we can see CO2 was the major GHG emitted by human beings. To observe the trend of change more clearly, graphs below show the concentration over years for each GHG. It is hard to summarize the general trend for all GHG emissions, so in the further analysis, the GHG type comparison will be first studied. 

```{r warning=FALSE}
#Show the structure of the raw data
str(GHG.gas)

#Rename column name for easier use
colnames(GHG.gas)
GHG.gas <- GHG.gas %>%
  rename("CO2" = Carbon.dioxide, 'CH4' = Methane, 'N2O' = Nitrous.oxide, 'Fgas' = Fluorinated.gases)
colnames(GHG.gas)

```

```{r echo = FALSE, message=FALSE, warning=FALSE, fig.cap= 'US GHG Concentration by Gas Type in 1990-2017 (MMTCO2e)', fig.height=8, fig.width=7}
#Preview of reframed dataset
kable(head(GHG.gas), caption = "First 6 years of GHG emissions by gas type (MMTCO2e)") %>%
  kable_styling(latex_options="hold_position")

#Gather dataset together for visualization
GHG.gas.gather <- gather(GHG.gas, 'Gas', 'Concentration', -Year)
count(GHG.gas.gather)

GHG.gas.line <- ggplot(GHG.gas.gather)+
  geom_line(aes(x= Year, y=Concentration, color = Gas))+
  labs(y= bquote(bold('Concentration (MMT'~CO[2]~'e)')),
       color = '')+
  scale_color_brewer(palette = 'Set1')+
  scale_x_continuous("Year", labels = as.character(GHG.gas.gather$Year), breaks = GHG.gas.gather$Year)

GHG.gas.line.facet <- ggplot(GHG.gas.gather)+
  geom_line(aes(x= Year, y=Concentration, color = Gas))+
  labs(y= bquote(bold('Concentration (MMT'~CO[2]~'e)')),
       color = '')+
  scale_color_brewer(palette = 'Set1')+
  facet_wrap(vars(Gas), scales = 'free')+
  theme(legend.position = "none")

plot_grid(GHG.gas.line,
          GHG.gas.line.facet,
          label_x = 0.2,
          nrow = 2)
```

\newpage

##US GHG Emissions by Sector
Figure 4 illustrates the GHG emissions over the same period by sector. The ideology of Figure 4 is same as Figure 3, which internally compares the emissions by sectors. The total sectoral GHG emission is same as the sum of GHG concentration by gas type. Among all 6 sectors emitting GHGs, electricity generation, transporation, and industial sector had highest emissions. It is inspired to study how sectoral emissions had contributed to total GHG emissions, and how they were different from each other.

```{r warning=FALSE}
str(GHG.sector)

#Rename column name for easier use
colnames(GHG.sector)
GHG.sector <- GHG.sector %>%
  rename('Electricity_generation' = Electricity.generation)
colnames(GHG.sector)

##Reframe dataset to meet project objective
GHG.sector.select <- GHG.sector %>%
  select(Year, Transportation, Electricity_generation, Industry, Agriculture, Commercial,Residential)
dim(GHG.sector.select)

```

```{r echo = FALSE, message=FALSE, warning=FALSE}
#Preview of reframed dataset
kable(head(GHG.sector.select), caption = "First 6 years of GHG emissions by sector (MMTCO2e)") %>%
  kable_styling(latex_options="hold_position")
```

```{r echo = FALSE, message=FALSE, warning=FALSE, fig.cap= 'US GHG Emissions by Sector in 1990-2017 (MMTCO2e)', fig.height=8, fig.width=7}

#Gather dataset together for visualization
GHG.sector.gather <- gather(GHG.sector.select, 'Sector', 'Emission', -Year)
count(GHG.sector.gather)

GHG.sector.line <- ggplot(GHG.sector.gather)+
  geom_line(aes(x= Year, y=Emission, color = Sector))+
  labs(y= bquote(bold('GHG Emission (MMT'~CO[2]~'e)')),
       color = '')+
  scale_color_brewer(palette = 'Dark2')+
  scale_x_continuous("Year", labels = as.character(GHG.sector.gather$Year), breaks = GHG.sector.gather$Year)

GHG.sector.line.facet <- ggplot(GHG.sector.gather)+
  geom_line(aes(x= Year, y=Emission, color = Sector))+
  labs(y= bquote(bold('GHG Emission (MMT'~CO[2]~'e)')),
       color = '')+
  scale_color_brewer(palette = 'Dark2')+
  facet_wrap(vars(Sector), scales = 'free')+
  theme(legend.position = "none")

plot_grid(GHG.sector.line,
          GHG.sector.line.facet,
          label_x = 0.2,
          nrow = 2)
```


\newpage

##US Mean Annual Temperature

Mean annual temperature and temperature anomalies over 1990-2017 are visualized in Figure 5. Temperature anomaly refers to the difference between annual mean temperature and overall averaage temperature for that period. Therefore, the overall average temperature were also calculated and mutated as a new column in the table. Since overall average temperature is constant, temperature anomalies are in the same changing trends as annual mean temperature. Since temperature anomalies are referene values, they can better and more easily demonstrate increasing and decreasing trend by positive and negative symbols. Therefore, temperature anomaly will be used in the future analyses.
```{r warning=FALSE}
str(temp)
#Convert year characters to date format
colnames(temp)
temp <- mutate(temp, Year = year(as.Date(as.yearmon(as.character(temp$Date), format = "%Y%m"))))
str(temp)

#Reframe dataset to meet project objective
temp.select <- temp %>%
  select(Year, Value, Anomaly) %>%
  mutate(Mean_all_temp =  mean(Value))
dim(temp.select)
str(temp.select)

```

```{r echo = FALSE, message=FALSE, warning=FALSE}
#Preview of reframed dataset
kable(head(temp.select), caption = "First 6 years of US temperature data in 1990-2017 (F)") %>%
  kable_styling(latex_options="hold_position")

```

```{r echo = FALSE, message=FALSE, warning=FALSE,fig.cap= "US Mean Annual Temperature Anomaly in 1990-2017 (F)"}
temp.line <- ggplot(temp.select, aes(x= Year, y=Value))+
  geom_point()+
  geom_line()+
  geom_line(aes(y=Mean_all_temp, color= 'red'))+
  labs(y= expression("Temperature ("*~degree*F*")"),
       title = expression("US Mean Annual Temperature in 1990-2017 ("*~degree*F*")"))+
  scale_x_continuous("Year", labels = as.character(temp.select$Year), breaks = temp.select$Year)+
  geom_text(x = 2017, y = 53, label = "Average Temperature for 18 Years", hjust = 1, size = 3, color = 'red')+
  theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_blank())

temp.select$colour <- ifelse(temp.select$Anomaly < 0, "negative","positive")
temp.anom <- ggplot(temp.select)+
  geom_col(aes(x=Year, y=Anomaly, fill=colour))+
  labs(y= expression("Temperature Anomaly ("*~degree*F*")"),
      color = '')+
  scale_x_continuous("Year", labels = as.character(temp.select$Year), breaks = temp.select$Year)+
  scale_fill_manual(values=c(positive="firebrick1",negative="steelblue"))+
  theme(legend.position = 'none')

plot_grid(temp.line,
          temp.anom,
          label_x = 0.2,
          nrow = 2)
```

\newpage

##US Population

Population data retrieved from World Bank is at global scale covering 60 years since 1960. Therefore, both geograhic and temporal selection are needed to reframe the dataset. The sample of the wrangled dateset is shown as Table ???.
```{r warning=FALSE}
str(ppl)

#Reframe dataset to meet project objective
ppl.select <- ppl %>%
  select(Year, United.States) %>%
  filter(Year >1989 & Year < 2018)
dim(ppl.select)

#Rename column name for easier use
colnames(ppl.select)
ppl.select <- ppl.select %>%
  rename('US_pop' = United.States)
colnames(ppl.select)
```

```{r echo = FALSE, message=FALSE, warning=FALSE}
#Preview of reframed dataset
kable(head(ppl.select), caption = "First 6 years of US population in 1990-2017") %>%
  kable_styling(latex_options="hold_position")
```

```{r message = FALSE, warning = FALSE}
#Save all processed code to the Processed folder
write.csv(GDP.select, row.names = FALSE, file = './Processed Data/BEA_GDP_processed.csv')
write.csv(elec.select, row.names = FALSE, file = './Processed Data/EIA_electricity-consumption_sector_processed.csv')
write.csv(GHG.gas, row.names = FALSE, file = './Processed Data/EPA_GHG_Gas_processed.csv')
write.csv(GHG.sector.select, row.names = FALSE, file = './Processed Data/EPA_GHG_Sector_processed.csv')
write.csv(temp.select, row.names = FALSE, file = './Processed Data/NOAA_temp_processed.csv')
write.csv(ppl.select, row.names = FALSE, file = './Processed Data/WB_pop_processed.csv')

```

```{r message = FALSE, warning = FALSE}

#Reload new datasets.
newGDP <- read.csv('./Processed Data/BEA_GDP_processed.csv', head(T))
newelectricity <- read.csv('./Processed Data/EIA_electricity-consumption_sector_processed.csv', head(T))
newGHGgas <- read.csv('./Processed Data/EPA_GHG_Gas_processed.csv', head(T))
newGHGsector <- read.csv('./Processed Data/EPA_GHG_Sector_processed.csv', head(T))
newtemp <- read.csv('./Processed Data/NOAA_temp_processed.csv', head(T))
newpop <- read.csv('./Processed Data/WB_pop_processed.csv', head(T))

allGHG <- full_join(newGHGgas,newGHGsector, by = 'Year')
GHG_contribute <- full_join(newelectricity, newpop, by = 'Year')
GHG_impact <- full_join(newGDP, newtemp, by = 'Year')
```

\newpage

#Analysis
*Figures in this sections are disordered - the float package was not compatible to my system. All statements in this section will describe which specific figures are referring to in certain analyses.*

##Question 1: What factors have contributed to GHG emissions in these 18 years?

###Are emissions of different GHGs related to each other?
As discussed in Chapter 3, there is no general trend for the concentration of all GHGs. Therefore, before delving into external impacts on GHG concentrations, how different types of GHGs were varied from each other were first studied.

The distribution of all GHG data were assessed with shapiro.test. Most p-values are  greater than 0.05, which implies that the distribution of these data re not significantly different from normal distribution. Therefore, these data are normally distributed. Other data, namely CH4 and Fgas, obtain a p-value less than 0.05 (even transforming these two variables by taking the log and square root), which indicate they are not normally distributed. However, from the normal Q-Q plot, the scatters lie close to the line and symmetry across the qq normal line, so these two data can still be seen as normally distributed. All GHG emissions by sector obtain p-values greater than 0.05 from the shapiro.test, so they are normally distributed. Parametric tests will be used for all GHG dataset.

One-sample t-test was first conducted to assess the null hypothesis that the mean of allGHG dataset is equal to the mean of total . For example, we might ask ourselves (from the EPA air quality processed dataset): 

Are Ozone levels below the threshold for "good" AQI index (0-50)?
lm - look at explainations here https://feliperego.github.io/blog/2015/10/23/Interpreting-Model-Output-In-R

```{r include = FALSE}

summary(allGHG)

shapiro.test(allGHG$Total)
shapiro.test(allGHG$CO2)
shapiro.test(allGHG$CH4)
shapiro.test(allGHG$N2O)
shapiro.test(allGHG$Fgas)
shapiro.test(sqrt(allGHG$CH4))
shapiro.test(sqrt(allGHG$Fgas))
qqnorm(allGHG$CH4);qqline(allGHG$CH4)


shapiro.test(allGHG$Transportation)
shapiro.test(allGHG$Electricity_generation)
shapiro.test(allGHG$Industry)
shapiro.test(allGHG$Agriculture)
shapiro.test(allGHG$Commercial)
shapiro.test(allGHG$Residential)

allsample <- t.test(allGHG, mu=mean(allGHG$Total), alternative = 'less')
allsample

```

lm: a small p-value for the intercept and the slope indicates that we can reject the null hypothesis which allows us to conclude that there is a relationship between speed and distance.
```{r include = FALSE}
GHG.gas.regression <- lm(data=allGHG, CO2~ CH4+N2O+Fgas)
summary(GHG.gas.regression)
step(GHG.gas.regression)

GHG.gas.regression.improve <- lm(formula = CO2 ~ CH4 + N2O, data = allGHG)
summary(GHG.gas.regression.improve)

allGHG.bygas <- allGHG %>%
  select(CO2:Fgas)

cor.test(allGHG.bygas$CO2, allGHG.bygas$CH4)
cor.test(allGHG.bygas$CO2, allGHG.bygas$N2O)
cor.test(allGHG.bygas$CO2, allGHG.bygas$Fgas)
#Visualize this result
allGHG.gas.corr <- cor(allGHG.bygas)
```

```{r echo = FALSE, message=FALSE, warning=FALSE, fig.cap= "Checking Model Fit and Assumptions for US GHG Concentration by Gas"}
par(mfrow = c(2,2), mar=c(1,1,1,1))
plot(GHG.gas.regression.improve)
par(mfrow = c(1,1))
```

```{r echo = FALSE, message=FALSE, warning=FALSE, fig.cap= "Correlation Plot of US GHG Concentration by Gas"}
corrplot(allGHG.gas.corr, method = "ellipse", tl.cex = 0.5)
```

\newpage

###Are sectoral GHG emissions significantly different from each other? How sectoral emissions contribute to total emissions (correlation)? (Lec12 - multiple regression)
cor.test: If the p-value is < 5%, then the correlation between x and y is significant.
```{r include = FALSE}
##sectoral emissions differ from each other?
GHG.regression <- lm(data=allGHG, Transportation~ Electricity_generation+Industry+Agriculture+Commercial+Residential)
summary(GHG.regression)
step(GHG.regression)

#how contribute to total emission - correlation
cor.test(allGHG$Total,allGHG$Transportation)
cor.test(allGHG$Total,allGHG$Electricity_generation)
cor.test(allGHG$Total,allGHG$Industry)
cor.test(allGHG$Total,allGHG$Agriculture)
cor.test(allGHG$Total,allGHG$Commercial)
cor.test(allGHG$Total,allGHG$Residential)

```

```{r echo = FALSE, message=FALSE, warning=FALSE, fig.cap= "Checking Model Fit and Assumptions for US GHG Emission by Sector"}
par(mfrow = c(2,2), mar=c(1,1,1,1))
plot(GHG.regression)
```

```{r include = FALSE}
par(mfrow = c(1,1))
allGHG.bysector <- allGHG %>%
  select(Total:Residential)
allGHG.corr <- cor(allGHG.bysector)
```

```{r echo = FALSE, message=FALSE, warning=FALSE, fig.cap= "Correlation Plot of US GHG Emission by Sector"}
par(mfrow = c(1,1))
corrplot(allGHG.corr, method = "ellipse", tl.cex = 0.5)
```

\newpage

###Check dataset- Does total electricity consumption differ among sectors (category vs discrete)?
Shapiro-Wilk test can be performed as follow:
Null hypothesis: the data are normally distributed
Alternative hypothesis: the data are not normally distributed

lm: If the p-value is less than 0.05, we reject the null hypothesis that there's no difference between the means and conclude that a significant difference does exist.
```{r include = FALSE}
newelectricity.gather <- gather(newelectricity, 'Sector', 'Emission', -Year)
count(newelectricity.gather)

newelectricity.gather$Emission <- str_replace_all(newelectricity.gather$Emission,
pattern = "([,])", replacement = "")
newelectricity.gather$Emission <- as.numeric(newelectricity.gather$Emission)
str(newelectricity.gather)
#elec.gather2 <- elec.gather %>%
 # filter(!is.na(Electricity_Consumption))

#One-way Anova, test for normality
shapiro.test(newelectricity.gather$Emission[newelectricity.gather$Sector=='elec_residential'])
shapiro.test(newelectricity.gather$Emission[newelectricity.gather$Sector=='elec_commercial'])
shapiro.test(newelectricity.gather$Emission[newelectricity.gather$Sector=='elec_industrial'])
shapiro.test(newelectricity.gather$Emission[newelectricity.gather$Sector=='elec_transport'])

#Test for equal variance
##a small p-value for the intercept and the slope indicates that we can reject the null hypothesis which allows us to conclude that there is a relationship between speed and distance.
bartlett.test(newelectricity.gather$Emission ~ newelectricity.gather$Sector)

newelectricity.anova <- aov(data = newelectricity.gather, Emission ~ Sector)
summary(newelectricity.anova)

```

```{r echo = FALSE, message=FALSE, warning=FALSE, fig.cap= "Checking Model Fit and Assumptions for US Electricity Consumption"}
#Checking model fit and assumptions
# ANOVA is robust against departures from normality.
par(mfrow = c(2,2), mar=c(1,1,1,1))
plot(newelectricity.anova)

```

```{r include = FALSE}

# Post-hoc test # Run a post-hoc test for pairwise differences
TukeyHSD(newelectricity.anova)

#Extract groupings for pairwise relationships
newelectricity.groups <- HSD.test(newelectricity.anova, "Sector", group = TRUE)
newelectricity.groups

```

```{r echo = FALSE, message=FALSE, warning=FALSE, fig.cap= "Post-hoc Test Plot - Pairwise Relationship between Emissions and Sectors"}

newelectricity.graph <- ggplot(newelectricity.gather, aes(x = Sector, y = Emission)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + stat_summary(geom = "text", fun.y = max, vjust = -1, size = 3.5, label = c('b','c','a','d')) +
  labs(x = "Sector", y = "Emission") 
print(newelectricity.graph)

```

\newpage

### Are electricity comsupmtions in some sectors significant to total GHG emissions?
lm: If the p-value is less than 0.05, we reject the null hypothesis which allows us to conclude that there is a relationship between speed and distance.
```{r include = FALSE}
##linear regression - electricity consumption relation to total GHG emissions
GHG_contribute$elec_residential <- str_replace_all(GHG_contribute$elec_residential,
pattern = "([,])", replacement = "")
GHG_contribute$elec_residential <- as.numeric(GHG_contribute$elec_residential)
GHG_contribute$elec_commercial <- str_replace_all(GHG_contribute$elec_commercial,
pattern = "([,])", replacement = "")
GHG_contribute$elec_commercial <- as.numeric(GHG_contribute$elec_commercial)
GHG_contribute$elec_industrial <- str_replace_all(GHG_contribute$elec_industrial,
pattern = "([,])", replacement = "")
GHG_contribute$elec_industrial <- as.numeric(GHG_contribute$elec_industrial)
GHG_contribute$elec_transport <- str_replace_all(GHG_contribute$elec_transport,
pattern = "([,])", replacement = "")
GHG_contribute$elec_transport <- as.numeric(GHG_contribute$elec_transport)

elec.regression <- lm(allGHG$Total ~GHG_contribute$elec_residential + GHG_contribute$elec_commercial + GHG_contribute$elec_industrial + GHG_contribute$elec_transport)
summary(elec.regression)

step(elec.regression)

elecmodel <- lm(formula = allGHG$Total ~ GHG_contribute$elec_residential + 
    GHG_contribute$elec_transport)
summary(elecmodel)
```

```{r echo = FALSE, message=FALSE, warning=FALSE, fig.cap= "Checking Model Fit and Assumptions for the Improved Model"}

par(mfrow = c(2,2), mar=c(1,1,1,1))
plot(elecmodel)
par(mfrow = c(1,1))
```

\newpage

###Is population growth significant to GHG emissions ?
cor.test: If the p-value is < 5%, then the correlation between x and y is significant.

```{r include = FALSE}
shapiro.test(GHG_contribute$US_pop) #normal

pop.regression <- lm(allGHG$Total ~GHG_contribute$US_pop)
summary(pop.regression)
cor.test(allGHG$Total,GHG_contribute$US_pop) #not significant, no relationship, same as lm
```

```{r echo = FALSE, message=FALSE, warning=FALSE, fig.cap= "Checking Model Fit and Assumptions for US Population"}

par(mfrow = c(2,2), mar=c(1,1,1,1))
plot(pop.regression)
par(mfrow = c(1,1))
```

```{r echo = FALSE, message=FALSE, warning=FALSE, fig.cap= "Relationship between Population and GHG Concentration"}

ppl.all <- full_join(allGHG, newpop, by = 'Year')
# Plot the regression
GHGpop.plot <-ggplot(ppl.all, aes(x = US_pop, y = Total)) + 
  geom_line()+
  labs(y= bquote(bold('GHG Emission (MMT'~CO[2]~'e)')),
       x= 'Population',
       color = '')+
  geom_vline(xintercept = ppl.all$US_pop[which.max(ppl.all$Total)], color = "#f20505", lty = 2, size = 0.9)+
  geom_text(y = 6400, x = 3.0e+08, label = 'Highest GHG Concentration Year: 2007', hjust = 1, size = 3, color = 'red')
print(GHGpop.plot)

```

\newpage

##Question 2: 
### GHGs and temperature 
which sector emission is significant to temperature anomaly?
```{r include = FALSE}
shapiro.test(newtemp$Anomaly)
#time series analysis- Mann-Kendall
#The null hypothesis of the test is that there is no trend in the data and the alternative hypothesis is that the data represents a monotonic trend.

#If the p-value is less than 0.05, we reject the null hypothesis.
GHGtemp_data<- full_join(allGHG, newtemp, by='Year')
GHGtemp.ts <- ts(GHGtemp_data$Value, start=c(1990), end=c(2017), frequency=1)
class(GHGtemp.ts)
GHGtemp.ts.vector <- as.numeric(GHGtemp.ts)
class(GHGtemp.ts.vector)
GHGtemp.trend <- mmkh(GHGtemp.ts.vector, ci=0.95)
GHGtemp.trend
summary(GHGtemp.trend)


#linear regression
##a p-value of 5% or less is a good cut-off point. In our model example, the p-values are very close to zero. Note the ‘signif. Codes’ associated to each estimate. Three stars (or asterisks) represent a highly significant p-value. Consequently, a small p-value for the intercept and the slope indicates that we can reject the null hypothesis which allows us to conclude that there is a relationship between speed and distance.
GHG.temp.regression <- lm(GHGtemp_data,formula = Anomaly~  Transportation+Electricity_generation+Industry+Agriculture+Commercial+Residential)
summary(GHG.temp.regression)

step(GHG.temp.regression)

GHG.temp.regression.improve <- lm(formula = Anomaly ~ Transportation + Agriculture + Commercial + Residential, data = GHGtemp_data)
summary(GHG.temp.regression.improve)


GHGsector.temp <- GHGtemp_data %>%
  select(Transportation:Residential, Anomaly)

cor.test(GHGsector.temp$Anomaly, GHGsector.temp$Transportation)
cor.test(GHGsector.temp$Anomaly, GHGsector.temp$Electricity_generation)
cor.test(GHGsector.temp$Anomaly, GHGsector.temp$Industry)
cor.test(GHGsector.temp$Anomaly, GHGsector.temp$Agriculture)
cor.test(GHGsector.temp$Anomaly, GHGsector.temp$Commercial)
cor.test(GHGsector.temp$Anomaly, GHGsector.temp$Residential)

GHGsector.temp.corr <- cor(GHGsector.temp)
```

```{r echo = FALSE, message=FALSE, warning=FALSE, fig.cap= "Checking Model Fit and Assumptions for the Relationship between US Mean Annual Temperature and Sectoral Emissions"}
par(mfrow = c(2,2), mar=c(1,1,1,1))
plot(GHG.temp.regression.improve)
par(mfrow = c(1,1))
```

```{r echo = FALSE, message=FALSE, warning=FALSE, fig.cap= "Correlation Plot between US Mean Annual Temperature and Sectoral GHG Emissions"}
corrplot(GHGsector.temp.corr, method = "ellipse", tl.cex = 0.5)

```

### GHG and GDP
Is GHG emission significant to GDP growth?
```{r include = FALSE}
shapiro.test(newGDP$GDP)

GHGgdp_data<- full_join(allGHG, newGDP, by='Year')
GHGgdp.ts <- ts(GHGgdp_data$GDP, start=c(1990), end=c(2017), frequency=1)
class(GHGgdp.ts)
GHGgdp.ts.vector <- as.numeric(GHGgdp.ts)
class(GHGgdp.ts.vector)
GHGgdp.trend <- mmkh(GHGgdp.ts.vector, ci=0.95)
GHGgdp.trend
summary(GHGtemp.trend)

GDP.anova <- lm(data=GHGgdp_data, Total~GDP+Personal_consumption)
summary(GDP.anova) 
step(GDP.anova)
```

```{r echo = FALSE, message=FALSE, warning=FALSE, fig.cap= "Checking Model Fit and Assumptions for the Relationship between US GDP and Total GHG Emissions"}
par(mfrow = c(2,2), mar=c(1,1,1,1))
plot(GDP.anova)
par(mfrow = c(1,1))
```

```{r echo = FALSE, message=FALSE, warning=FALSE, fig.cap= "Relationship between US GDP and Total GHG Emissions"}
# Add a line and standard error for the linear regression
GHGgdp.plot <- ggplot(GHGgdp_data, aes(x = Total, y = GDP)) + 
  geom_smooth(method = "lm") +
  labs(y= 'GDP (US$ billion)', x= bquote(bold('Total GHG Emission (MMT'~CO[2]~'e)')))+
  geom_point()
print(GHGgdp.plot)
```

\newpage

# Summary and Conclusions


\newpage

# References
<add references here if relevant, otherwise delete this section> 
